Towards Practical First-Order Model Counting

### 2024-07-17 doc/paper/paper.tex
A wrapper for FastWFOMC.jl

@@

  $\mathcal{B} \gets \emptyset$\;
  \ForEach{equation $(f(\mathbf{x}) = \expr{}) \in \mathcal{E}$}{
    \ForEach{function call $f(\mathbf{y}) \in \expr{}$}{\label{line:functioncall}
      \ForEach{$y_{i} \in \mathbf{y}$} {
        \uIf{$y_{i} \in \mathbb{N}_{0}$}{
          $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto y_{i}] \,\}$\;
        }
        \ElseIf{$y_{i} = x_{i} - c_{i}$ for some $c_{i} \in \mathbb{N}_{0}$}{
          \For{$j \gets 0$ \KwTo $c_{i} - 1$}{\label{line:lim}
            $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto j] \,\}$\;\label{line:insert}
          }
        }
      }
    }
  }
\end{algorithm}
% TODO: replace x_i-c_i on Line 7 of Algorithm 2 with a generalisation of x_i
% that can be an arbitrary summation/subtraction of domain sizes (propagating
% this change to the rest of the paper)

% TODO: a function call with a constant should always mean that that function
% should have a base case, not just when it's in the same function

@@

\begin{example}
  Let us consider the recursive function $g$ from \cref{example:overall}. In
  this case, \FindBaseCases iterates over two function calls: $g(l-1, m)$ and
  $g(l-1, m-1)$. The former produces the base case $g(0, m)$, while the latter
  produces both $g(0, m)$ and $g(l, 0)$.
\end{example}

@@

Finally, we assume a particular order of evaluation for function calls using the
equations in $\mathcal{E}$. Specifically, we assume that base cases are
considered before the recursive definition. The exact order in which base cases
are considered is not relevant.

@@

With the facts and assumptions mentioned above, we prove the following theorem.

@@

  Otherwise, let $f(y)$ be an arbitrary function call on the right-hand side of
  the definition of $f(x)$. If $y$ is a constant, there is a base case for
  $f(y)$. Otherwise, let $y = x - c$ for some $c > 0$. Then there exists
  $k \in \mathbb{N}_{0}$ such that $0 \le x - kc \le c-1$. So, after $k$
  iterations, the sequence of function calls $f(x), f(x-c), f(x-2c),\dots$ will
  be captured by the base case $f(x \mod c)$.
\end{proof}

@@

\section{Experimental Evaluation}\label{sec:experiments}
% TODO: must rerun on an HPC and include more instances

% TODO: this section is apparently "hard to follow"

@@

% TODO: maybe instead of a table I should have an experiment where:
% 1) Crane scales exponentially (with a base of sqrt(2))
% 2) other algorithm scale linearly
% 3) Then I can just draw the usual cumulative plot
% 4) Stop each algorithm when they reach the timeout value

@@

% TODO: I'm claiming that C++ code generation is not the main contribution, but
% the experiments are mainly about it...

% TODO: The asymptotic analysis of the total number of arithmetic operations
% required to compute the FOMC of these formulas was already performed in
% previous work [1].

% TODO: Since FastWFOMC does not support many-sorted logic, our experiments were
% restricted to formulas with a single domain. However, many of the counting
% problems examined in [1] become equivalent (in the sense that they generate
% the same integer sequences) when restricted to a single domain. Moreover, it
% is challenging to compare Crane2 and FastWFOMC on a larger set of problems
% because each problem has to be represented in two (fairly different) logics:
% FO and UFO^2+EQ.

% TODO: position some of the experiments as 'practical'

>>>>

% TODO: (somewhere) The implementation is only for unweighted (but the
% definition of WFOMC is still necessary because of Skolemization).

@@

  $\mathcal{B} \gets \emptyset$\;
  \ForEach{function call $f(\mathbf{y})$ on the right-hand side of an equation in $\mathcal{E}$}{\label{line:functioncall}
    $\mathbf{x} \gets \text{the parameters of $f$ in its definition}$\;
    \ForEach{$y_{i} \in \mathbf{y}$} {
      \uIf{$y_{i} \in \mathbb{N}_{0}$}{
        $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto y_{i}] \,\}$\;
      }
      \ElseIf{$y_{i} = x_{i} - c_{i}$ for some $c_{i} \in \mathbb{N}_{0}$}{
        \For{$j \gets 0$ \KwTo $c_{i} - 1$}{\label{line:lim}
          $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto j] \,\}$\;\label{line:insert}
        }
      }
    }
  }
\end{algorithm}
% TODO (later, let's not bother yet): replace x_i-c_i on Line 7 of Algorithm 2
% with a generalisation of x_i that can be an arbitrary summation/subtraction of
% domain sizes (propagating this change to the rest of the paper, particularly
% Assumption 1, the proof of Lemma 1)

@@

\begin{example}
  Consider the recursive function $g$ from \cref{example:overall}. Then
  \FindBaseCases iterates over two function calls: $g(l-1, m)$ and
  $g(l-1, m-1)$. The former produces the base case $g(0, m)$, while the latter
  produces both $g(0, m)$ and $g(l, 0)$.
\end{example}

@@

Finally, we assume a particular order of evaluation for function calls using the
equations in $\mathcal{E}$. Specifically, we assume that base cases are
considered before the recursive definition. The exact order in which base cases
are considered is immaterial.

@@

With the observations and assumptions mentioned above, we prove the following
theorem.

@@

  Otherwise, let $f(y)$ be an arbitrary function call on the right-hand side of
  the definition of $f(x)$. If $y$ is a constant, then there is a base case for
  $f(y)$. Otherwise, let $y = x - c$ for some $c > 0$. Then there exists
  $k \in \mathbb{N}_{0}$ such that $0 \le x - kc \le c-1$. So, after $k$
  iterations, the sequence of function calls $f(x), f(x-c), f(x-2c),\dots$ will
  be captured by the base case $f(x \mod c)$.
\end{proof}

@@

\section{Experimental Evaluation}\label{sec:experiments}
% TODO: this section is apparently "hard to follow"

@@

@@

% TODO: The asymptotic analysis of the total number of arithmetic operations
% required to compute the FOMC of these formulas was already performed in
% previous work.

% TODO: Since FastWFOMC does not support many-sorted logic, our experiments were
% restricted to formulas with a single domain. However, many of the counting
% problems examined in [1] become equivalent (in the sense that they generate
% the same integer sequences) when restricted to a single domain. Moreover, it
% is challenging to compare Crane2 and FastWFOMC on a larger set of problems
% because each problem has to be represented in two (fairly different) logics:
% \FO{} and \UFO{}.
